---
title: "Analysis - Franklin Glance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How did the rise in technology usage (social media, internet, etc..) aid or hurt historically successful businesses/industries? 

Loading Packages
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(forcats)
library(lubridate)
library(sparklyr)
library(ggthemes)
library(gridExtra)
library(readxl)
library(stringr)
library(corrr)
library(here)
library(corrgram)
library(readr)
setwd("/Users/franklinglance/Library/CloudStorage/OneDrive-UniversityofVirginia/UVA/Spring 2022/SYS 2202/Final Project/Fortune_500_Analysis")
```

## Importing Data
```{r}
# us company data for 2021 Data was collected for the year end February 2021
fortune_1000 <- read_csv("data/Fortune_1000.csv") # provides sector info for selection of fortune 500 companies
fortune1955_2021 <- read_csv("data/fortune500_1955_2021.csv")  # rank + revenue data for fortune 500 companies
percent_using_internet <- read_excel("data/percent_of_population_using_internet.xlsx")
technology_adoption_by_households_in_the_united_states <- read_csv("data/technology-adoption-by-households-in-the-united-states.csv")
computing_efficiency <- read_csv("data/computing-efficiency.csv")
# data on various business indicators (see metadata)
data_bisiness_indicators <- read_csv("data/data_business_indicators/aee336d2-021b-4021-912c-3de7fd9d2729_Data.csv")
# data on supercomputer power
supercomputer_power_flops <- read_csv("data/supercomputer-power-flops.csv")

# data on phone subscriptions
landline_cellular_data <- read_excel("data/landline_cellular_data.xlsx")
```

## Tidying Data

Prepare Fortune 1000 and Fortune1995_2021 for joining + join by company
```{r}
fortune_1000 <- fortune_1000 %>%
  mutate(Company = company)
fortune1955_2021 <- fortune1955_2021 %>%
  mutate(Company = Name)
```


```{r}
fortune_joined <- fortune_1000 %>%
  left_join(fortune1955_2021, by = "Company") %>%
  drop_na(sector)
# drop irrelavent columns
fortune_joined <- fortune_joined %>% 
  select(-c(rank, rank_change, revenue, profit, `num. of employees`, profitable, prev_rank, `Market Cap`))
```

Tidying Fortune Joined Dataset

- Revenue is now in billions of dollars
```{r}
# changing revenue to numeric type, now in billions of dollars
fortune_joined <- fortune_joined %>%
  transform(Revenue = as.integer(Revenue)) %>%
  mutate(Revenue = Revenue/1000) %>%
  transform(Revenue = as.integer(Revenue))
```


## Determining which companies are "successful" vs. "unsuccessful

### Can rank and Revenue be used synonomously when talking about success?

It appears that there is absolutely a correlation between rank and revenue, especially when looking at companies ranked in the top 200. Therefore, we will choose to focus on top 200 ranked companies, since lower ranking companies do not appear to differ significantly in revenue. 
```{r}
fortune_joined %>%
  filter(Year > 1985) %>%
  group_by(Rank) %>%
  summarise(mean_revenue_by_rank = mean(Revenue, na.rm = TRUE)) %>%
  ggplot(aes(Rank, mean_revenue_by_rank)) +
  geom_smooth() + 
  labs(title = "Mean Revenue by Rank, Fortune 500 from 1985 to 2021", 
       subtitle = "Displays correlation between rank and revenue") +
  ylab("Mean Revenue (in billions)") + 
  scale_x_continuous(trans = "reverse")
```

Viewing top 20 companes that have been on this list the longest.
```{r}
head(fortune_joined %>% count(Company, sort=TRUE), n=20)
```
Listing companies which have been on the fortune 500 for more than 20 years, after 1985.
```{r}

head(fortune_joined %>% count(Company, sort=TRUE), n=20)

# filtering by companies on the list for 20 years, since 1985
fortune_joined %>% 
  group_by(Company) %>%
  filter(Year > 1985) %>%
  filter(n() > 20) %>%
  summarise(unique(Company), n()) %>%
  arrange(desc(n()))
```


Now we will filter for companies who have been on the fortune 500 the entire time
```{r}
fortune_joined %>% 
  group_by(Company) %>%
  filter(Year > 1985) %>%
  filter(n() > 35) 
```


We know that the top 100 companes typically show the greatest correlation between rank and revenue, so we will filter for companes which were in the top 100 in 2021, and have been on the list since 1985. They are the following:
```{r}
top_100 <- fortune_joined %>% 
  group_by(Company) %>%
  filter(Year > 1985) %>%
  filter(n() > 35) %>%
  group_by(Company) %>%
  filter(Year == 2021) %>%
  filter(Rank < 100) %>%
  summarise(Company)

```

Choosing a distinct set of several "successful" businesses will allow a more in depth analysis when analyzing correlation with technological advancements. 
```{r}
as.vector(top_100)

# creating vector of companies who have been in the top 100 the whole time
top_100_vector = c(t(top_100))
```



Creating data frame with top companies from 2021 over the entire timeframe. Good subset of "successful" businesses. 
```{r}
top_100_companies_long_term <- fortune_joined %>% 
  group_by(Company) %>%
  filter(Year > 1985) %>%
  filter(n() > 35) %>%
  filter(Company %in% top_100_vector)
```


This isn't a great visualization, however it does show that these companies as a whole experienced growth over the years. 
```{r}
top_100_companies_long_term %>%
  group_by(Company) %>%
  ggplot(aes(Year, Revenue)) +
  geom_col(aes(fill = Company)) +
  labs(title = "Total Revenue of top 100 Companies from 1985 to 2021", 
       subtitle = "Displays distribution of Revenue by company") +
  ylab("Revenue (in billions)")
```
Top 100 revenue by sector. This is a very useful visualization, as it shows how some sectors have grown (technology, aerospace and defense), and others have diminished (motor vehicles and parts)
```{r}
top_100_companies_long_term %>%
  group_by(Company) %>%
  ggplot(aes(Year, Revenue)) +
  geom_col(aes(fill = sector)) +
  labs(title = "Total Revenue of top 100 Companies from 1985 to 2021, by Sector", 
       subtitle = "Displays distribution of Revenue by sector") +
  ylab("Revenue (in billions)")
```
Let us visualize the same thing, but for all companies on the fortune 500 since 1985. 
```{r}
fortune_joined %>% 
  group_by(Company) %>%
  filter(Year > 1985) %>%
  filter(n() > 35) %>%
  ggplot(aes(Year, Revenue)) +
  geom_col(aes(fill = sector)) +
  labs(title = "Total Revenue of fortune 500 Companies from 1985 to 2021, by Sector", 
       subtitle = "Revenue by sector for companes that were on the Fortune 500 since 1985") +
  ylab("Revenue (in billions)")
```
These graphs are helpful, however i think one focusing on the sectors could provide more useful insights. 
```{r}
top_100_companies_long_term %>% 
  group_by(Company) %>%
  filter(Year > 1985) %>%
  filter(n() > 35) %>%
  group_by(Year, sector) %>%
  summarise(revenue_per_year = sum(Revenue)) %>%
  ggplot(aes(Year, revenue_per_year)) +
  geom_col() +
  facet_wrap(~ sector) +
  labs(title = "Total Revenue of top 100 Companies from 1985 to 2021, by Sector", 
       subtitle = "Displays distribution of Revenue by sector") +
  ylab("Revenue (in billions)")
```
```{r}
fortune_joined %>% 
  group_by(Company) %>%
  filter(Year > 1985) %>%
  filter(n() > 35) %>%
  group_by(Year, sector) %>%
  summarise(revenue_per_year = sum(Revenue)) %>%
  ggplot(aes(Year, revenue_per_year)) +
  geom_col() +
  facet_wrap(~ sector) +
  labs(title = "Total Revenue of fortune 500 companies from 1985 to 2021, by Sector", 
       subtitle = "Displays distribution of Revenue by sector") +
  ylab("Revenue (in billions)")
```

```{r}
fortune_joined %>%
  group_by(Year, sector) %>%
  ggplot(aes(Year)) +
  geom_bar(aes(fill = sector)) +
  labs(title = "Successful Businesses Sector Distribution Over the Years")
```

The previous analysis has aided in determining the following insights about fortune 500 companies. 
- There is a correlation between rank and revenue
- Good subset of Companies to analyze are the top 100 companies from 2021, those of which have been on the fortune 500 since 1985. These can be our subset of "successful businesses"
- We can visualize the breakdown of successful businesses by sector, which provides a good look into which sectors have grown/shrunk since 1985. 


### Alternate subset of successful businesses (new additions to fortune 500 since 2000). 

```{r}
new_additions <- fortune_joined %>% 
  group_by(Company) %>%
  filter(Year > 2000) %>%
  filter(n() > 20) %>%
  group_by(Company) %>%
  filter(Year == 2021) %>%
  summarise(Company)

```



## Effects of Technology on Successful Businesses

Now we will explore potential relationships between emerging technologies and successful businesses. 






#### Technology Adoption by Househods in the US. 
Insights:
  - 
Determining which are the top entities
```{r}
technology_adoption_by_households_in_the_united_states %>% 
  group_by(`Entity`) %>% 
  filter(Year > 1985) %>%
  summarise(mean_diffusion_by_entity = mean(`Technology Diffusion (Comin and Hobijn (2004) and others)` )) %>%
  arrange(desc(mean_diffusion_by_entity))
```

Graphing just the top entities
```{r}
technology_adoption_by_households_in_the_united_states %>% 
  group_by(`Entity`) %>% 
  filter(Year > 1985) %>%
  mutate(mean_diffusion_by_entity = mean(`Technology Diffusion (Comin and Hobijn (2004) and others)` )) %>%
  filter(mean_diffusion_by_entity > 92) %>%
  ggplot(aes(Year, `Technology Diffusion (Comin and Hobijn (2004) and others)`)) +
  geom_smooth(aes(linetype = `Entity`)) +
  ylab("Technology Adoption, in millions")
```


#### Supercomputer power, 1993 to 2021:
In itself, the doubling of transistors every two years does not directly matter as much as how those changes apply to computing power. 

Insights:
  - the period of most rapid change in supercomputer power occurred from 2012 to 2021. This could be a good time frame to analyze the growth of tech succsessful businesses vs other successful businesses. 
```{r}
supercomputer_power_flops %>%
  group_by(Year) %>%
  ggplot(aes(Year, `Floating-Point Operations per Second`)) +
  geom_smooth() +
  labs(title = "Supercomputer Power, 1993 to 2021", 
       subtitle = "Floating-Point Operations per Second") + 
  ylab("Supercomputer Power (FLOPS)")
```

#### Computing efficiency, 1985 to 2015:

The cost to keep the machine running also matters. Computing efficiency measures the computational capacity per unit of energy. 
In this chart we see the computing efficiency of various processors over time. Here, computing efficiency is measured as the number of watts (a measure of electrical power) needed to carry out a million instructions per second (Watts per MIPS).

Insights:
  - the period of most rapid change in computing efficiency was during the time frame of 1990 to 2000. This could be a good time frame to analyze the growth of tech succsessful businesses vs other successful businesses. 
```{r}
computing_efficiency %>%
  group_by(Year) %>%
  filter(Year > 1985) %>%
  ggplot(aes(Year, log10(`Computing efficiency`))) +
  geom_smooth() +
  labs(title = "Computing efficiency, 1985 to 2015", 
       subtitle = "Computer processing efficiency, Watts per MIPS.") + 
  ylab("Computing Efficiency")
```

### Internet Usage, 1990 - 2019
While computing efficiency and supercomputer power directly relate to the power of tech, the adoption of tech is also very important. To understand this, we will look at data regarding the percentage of people in the US using the internet

Insights:
  - 
```{r}
percent_using_internet %>%
  ggplot(aes(Year, `Individuals using the Internet (% of population)`)) +
  geom_smooth() + 
  labs(title ="Percent of US population using the Internet, 1990 - 2019") + ylab("% of Population")
```

### Phone Usage, 1985 - 2019
```{r}
landline_cellular_data %>%
  ggplot(aes(Year, `Landline Subscriptions (per 100 people)`)) +
  geom_smooth() + 
  labs(title = "Landline Subscriptions,1985-2019") + 
  ylab("Subscriptions per 100 people")
```
```{r}
landline_cellular_data %>%
  ggplot(aes(Year, `Cellular Subscriptions (per 100 people)`)) +
  geom_smooth() + 
  labs(title = "Cellular Subscriptions per 100 people,1985-2019") + 
  ylab("Subscriptions per 100 people")
```
```{r}
landline_cellular_data %>%
  ggplot(aes(Year, `Cellular Subscriptions`/1000000)) +
  geom_smooth() + 
  labs(title = "Total Cellular Subscriptions, 1985-2019") + 
  ylab("Subscriptions (in millions)")
```

## Analysis of Relationship between Business Success and Technological Advancements/Adoption

We have seen above how various technologies have grown and been adopted since 1985. We will now analyze the relationship betweent he growth of these technologies and the success rates of historically successful businesses. Additionally, we will analyze a subset of successful businesses, those of which that have joined the fortune 500 since 2000. 



### Relationship betweeen top 100 companies and tech advancements

joining datasets by year
```{r}
top_100_joined <- top_100_companies_long_term %>%
  left_join(landline_cellular_data, by = "Year") %>%
  left_join(percent_using_internet, by = "Year") %>%
  left_join(technology_adoption_by_households_in_the_united_states, by = "Year") %>%
  left_join(computing_efficiency, by = "Year") %>%
  left_join(supercomputer_power_flops, by = "Year")
```


All sectors
```{r}
top_100_joined %>%
  ggplot(aes(`Individuals using the Internet (% of population)`, Revenue)) + 
  geom_smooth() + 
  labs(title = "Relationship between % of US pop. using the Internet and Fortune 500 Revenue", 
       subtitle ="(top 100 selection of fortune 500)")
```
Tech sectors
```{r}
top_100_joined %>%
  filter(sector == "Technology") %>%
  ggplot(aes(`Individuals using the Internet (% of population)`, Revenue)) + 
  geom_smooth() +
    labs(title = "Relationship between % of US pop. using the Internet and Fortune 500 Revenue", 
       subtitle ="(top 100 selection of fortune 500, technology sector only)")
```

```{r}
top_100_joined %>%
  ggplot(aes(`Technology Diffusion (Comin and Hobijn (2004) and others)`, Revenue)) + 
  geom_smooth() +
    labs(title = "Relationship between technology diffusion and Fortune 500 Revenue", 
       subtitle ="(top 100 selection of fortune 500)")
```

```{r}
top_100_joined %>%
  ggplot(aes(`Computing efficiency`, Revenue)) + 
  geom_smooth() +
    labs(title = "Relationship between Computing Efficiency and Fortune 500 Revenue", 
       subtitle ="(top 100 selection of fortune 500)") +
    scale_x_continuous(trans = "reverse")
```
```{r}
top_100_joined %>%
  filter(sector == "Technology") %>%
  ggplot(aes(`Computing efficiency`, Revenue)) + 
  geom_smooth() +
    labs(title = "Relationship betwen Computing Efficiency and Revenue -> Technology Sector", 
       subtitle ="(top 100 selection of fortune 500, technology sector)")  +
    scale_x_continuous(trans = "reverse")
```


Technology as a sector experienced significantly faster growth after the US population surpassed 50% internet usage by capita. 



```{r}
top_100_joined %>%
  filter(sector == "Technology") %>%
  ggplot(aes(`Cellular Subscriptions`, Revenue)) + 
  geom_smooth() +
    labs(title = "Relationship betwen Cellular Subscriptions and Revenue -> Technology Sector", 
       subtitle ="(top 100 selection of fortune 500, technology sector)")  
```
```{r}
top_100_joined %>%
  ggplot(aes(`Cellular Subscriptions`, Revenue)) + 
  geom_smooth() +
    labs(title = "Relationship betwen Cellular Subscriptions and Revenue -> Top 100", 
       subtitle ="(top 100 selection of fortune 500, technology sector)")  
```










Correlation coefficients
```{r}
cor(top_100_joined$Revenue, top_100_joined$Rank, use = "complete.obs")


cor(top_100_joined$Revenue, top_100_joined$`Computing efficiency`, use = "complete.obs")
```

```{r}
corrgram(top_100_joined)
```




### Variable Breakdown (tentative)

Independent Variables:
- Internet Usage from 1990-2019
- Tech adoption by household 
- social network usage/reach among fortune 500 ceos (2012 - 2016)
- Broadband usage by household
- Sector (selecting by sector)

Dependent Variables:
- Sector, Revenue, Rank, of successful businesses. 
-> used to indicate relative "success" of a company over the years







